Examining the critical path neighborhood for flexible job shop problems

### 2016-01-25 paper/paper.tex
updated work on paper

\documentclass[a4paper,10pt]{article}
\usepackage{amsmath}
\usepackage{enumitem}

@@

\title{A local search approach for flexible job shop problems}
\author{Alex van Poppelen\smallskip\\
\emph{Supervisor:} Dr. J.A. Hoogeveen}

@@

\[ T_{i,j} \in \mathcal{T} = \{T_1, \dots, T_t\} \]

Each machine type $T_k$ has a given number of machines, defined as $m_k$. We will use $T_{k,l}$ to refer to the $l$-th machine of type $T_k$. Furthermore, each operation $O_{i,j}$ has a processing time $p_{i,j} > 0$ and a delay time $w_{i,j} \geq 0$.

@@

In some other literature describing the Flexible Job Shop problem, there is no concept of a machine type. Instead, each operation is defined to be processed on some subset of the set of all machines \cite{approxfjsp}. This formulation is more general than the one used in this paper.

@@

In some situations, it may be unrealistic to model a scheduling problem as a Flexible Job Shop instance. For example, a manufacturing scenario might introduce storage space constraints. The FJS assumes that a job finished processing on one machine can be removed and placed into an intermediate buffer immediately, while waiting for it to be scheduled onto the next machine. This assumption may be impractical, and a factory might not have room for any intermediate buffers.

@@

This definition introduces several new possible scenarios that must be resolved. For example, what happens when machine $T_{1,1}$ is blocked, waiting for machine $T_{2,1}$ to be freed, while $T_{2,1}$ is blocked waiting for $T_{1,1}$? This depends on the problem scenario, and the mechanisms in place to transfer jobs between machines. In this paper, we will assume that these mechanisms are capable of simultaneously swapping jobs between all blocked machines. Thus, any \emph{blocking cycles} will not cause a deadlock in the schedule.

@@

We decide to modify the definition of an undirected edge. Instead, it will define two operations that may be processed in any order on the same machine, or independently on two different machines of the same type. Directing these edges will define the order of processing in the case the operations are on the same machine. In the case that they aren't, it only means that the start time of the second operation is equal to or greater than the start time of the first operation. This will make the serialization of a schedule more difficult, but the advantage is that generating the neighborhood will be easier.

@@

Once all the edges in the disjunctive graph have been given directions, such that the graph is acyclic, it is possible to generate the schedule. The serialization function will assign operations to machines, and calculate the start time for each operation. The makespan can be determined after this process. It is important that the graph representation combined with the serialization function spans the entire set of feasible schedules. Furthermore, the serialization function should be deterministic in order to allow the local search to do its work.

To do this, we place the unscheduled operations in ordered queues for each machine type $T_k$. We take the first operation that has no unscheduled job predecessor, and schedule it greedily into the first available machine, taking care to schedule it at the same start time or after that of the previously scheduled operation for that type. This ensures the ordering on the operations is preserved in the schedule, and conforms to the requirements we laid out earlier in the representation section. Scheduling operations before the starting time of a previously scheduled operation on the same machine type can lead to a schedule that is actually infeasible.

@@

We could choose to apply several tricks to generate a better schedule, but instead we will let the local search do that for us by modifying the order of the operations for each machine type.

@@

\[ O_1 \rightarrow O_8 \]
\[ O_2 \rightarrow O_6 \rightarrow O_4 \]
\[ O_7 \rightarrow O_5 \rightarrow O_3 \]

@@

@@

For the FJS, we can define the longest path in a similar manner. Whereas the longest path for the classical Job Shop problem was also chronologically sequential in terms of the generated schedule, this is no longer necessarily the case for the FJS. We must also consider operations on the same machine type, scheduled on different machines. For example, consider the operations $O_{1,1}$ and $O_{2,2}$ from different jobs that must be scheduled on machines of type $T$, connected by a directed edge from $O_{1,1}$ to $O_{2,2}$. Suppose $O_{1,1}$ has been scheduled at time $t_1$ on machine $M_1$ of type $T$ due to a precedence constraint. Further suppose that machine $M_2$, also of type $T$, is available from some time $t_2 < t_1$. However, the serialization function is forced to schedule $O_{2,2}$ on $M_2$ at time $t_1$ anyway. In this scenario, we would like to reverse the order of the arc between $O_{1,1}$ and $O_{2,2}$, so that $O_{2,2}$ may be scheduled earlier and therefore potentially decrease the length of the longest path.

@@

Let's consider the absolute ordering on the set of operations on a machine type, as defined by the acyclic directed graph corresponding to a feasible schedule. In the classical Job Shop case, a critical block swap identifies two consecutively scheduled nodes on a machine. These two nodes are necessarily consecutive in the absolute ordering, as well. In the FJS problem, this is no longer necessarily the case, since operations in between may have been scheduled on other machines of the same type. If one of these other operations is of the same job as either of the selected nodes, simply reversing the directed edge between the two nodes will lead to one or more cycles.

@@

	\ganttgroup[inline=false]{$T_{1,1}$}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,12}{3}{4}
	\ganttbar[bar/.style={fill=blue}]{1,10}{6}{7} \\

	\ganttgroup[inline=false]{$T_{1,2}$}{0}{1}
	\ganttbar[bar/.style={fill=blue}]{1,9}{2}{5}
	\ganttbar[bar/.style={fill=yellow}]{2,13}{6}{8} \\

	\ganttgroup[inline=false]{$T_{2,1}$}{0}{1}
	\ganttbar[bar/.style={fill=blue}]{1,8}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,14}{9}{11} \\

	\ganttgroup[inline=false]{$T_{3,1}$}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,11}{0}{2}

@@

Consider the schedule shown in \textbf{Figure \ref{fig:sch-swap}}. The absolute ordering that generated this schedule follows:

@@

We wish to swap the order of operations $O_{1,9}$ and $O_{2,13}$, both on the critical path. Simply swapping their places leads to cycles, since the job operations will be out of order. We can resolve this situation by reversing a few more arcs, to achieve the following absolute ordering.

@@

Suppose the method above was not able to swap two operations, $O_{i,a_1}$ and $O_{j,b_2}$ for $i \neq j$. Then there exist operations $O_{i,a_2}$ and $O_{j,b_1}$ with $a_1 < a_2$ and $b_1 < b_2$ such that the absolute order is equivalent to the following:

@@

The critical path neighborhood for the classical Job Shop is guaranteed to consist of only feasible schedules, a very nice result from Van Laarhoven et al \cite{simulatedannealingjsp}. Unfortunately, this is not the case for the Flexible Job Shop. We have slightly modified the swap operation to ensure no cycles within a machine type are created. However, this does not guarantee no cycles in the entire graph are generated. Consider the schedule shown in \textbf{Figure \ref{fig:sch-cycle}}, where colors correspond to jobs.

@@

	\ganttgroup[inline=false]{$T_{1,1}$}{656}{657}
	\ganttbar[bar/.style={fill=blue}]{1225}{656}{664}
	\ganttbar[bar/.style={fill=green}]{1298}{665}{668} \\

	\ganttgroup[inline=false]{$T_{1,2}$}{656}{657}
	\ganttbar[bar/.style={fill=yellow}]{905}{657}{665} \\

	\ganttgroup[inline=false]{$T_{1,3}$}{656}{657}
	\ganttbar[bar/.style={fill=darkblue}]{1009}{657}{658}
	\ganttbar[bar/.style={fill=red}]{1055}{659}{661}
	\ganttbar[bar/.style={fill=orange}]{1249}{662}{668} \\

@@

	\ganttgroup[inline=false]{$T_{4,1}$}{656}{657}
	\ganttbar[bar/.style={fill=darkblue}]{1010}{660}{663} \\

	\ganttgroup[inline=false]{$T_{4,4}$}{656}{657}
	\ganttbar[bar/.style={fill=green}]{1297}{660}{660}

@@

The acyclic directed graph that generated this schedule could have the following partial absolute ordering on machine type $T_1$:

@@

@@

\subsection{Changes}

Intuitively, we can use the exact same representation and neighborhood for the FBJS problem as for the FJS.

%\addcontentsline{toc}{section}{References}
\printbibliography[heading=bibintoc]

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{booktabs}

@@

\title{Examining the critical path neighborhood for flexible job shop problems}
\author{Alex van Poppelen\smallskip\\
\emph{Supervisor:} Dr. J.A. Hoogeveen}

@@

\[ \mathcal{T} = \{T_1, \dots, T_t\} \]

Furthermore, each operation $O_{i,j}$ has a processing time $p_{i,j} > 0$ and a delay time $w_{i,j} \geq 0$. Each machine type $T_k$ has a given number of machines, defined as $m_k$. We will use $M_{k,l}$ to refer to the $l$-th machine of type $T_k$.

@@

In some other literature describing the Flexible Job Shop problem, there is no concept of a machine type. Instead, each operation is defined to be processed on some subset of the set of all machines \cite{approxfjsp, ls-fbjs}. This formulation is more general than the one used in this paper.

@@

In some situations, it may be unrealistic to model a scheduling problem as a Flexible Job Shop instance. For example, a manufacturing scenario might introduce storage space constraints. The FJS assumes that a job finished processing on one machine can be removed and placed into an intermediate buffer immediately, while waiting for the next machine to become available. This assumption may be impractical, and a factory might not have room for any intermediate buffers.

@@

This definition introduces a new questions that must be resolved. For example, what happens when machine $M_{1,1}$ is blocked, waiting for machine $M_{2,1}$ to be freed, while $M_{2,1}$ is blocked waiting for $M_{1,1}$? This depends on the problem scenario, and the mechanisms in place to transfer jobs between machines. In this paper, we will assume that these mechanisms are capable of simultaneously swapping jobs between all blocked machines. Thus, any \emph{pure blocking cycles} will not cause a deadlock in the schedule.

@@

We decide to modify the definition of an undirected edge. Instead, it will define two operations that may be processed in any order on the same machine, or independently on two different machines of the same type. Directing these edges will define the order of processing in the case the operations are on the same machine. In the case that they aren't, it only means that the start time of the second operation is equal to or greater than the start time of the first operation. This will make the serialization of a schedule more difficult, but the advantage is that generating the neighborhood will remain easy.

@@

To generate a solution, all the edges in the disjunctive graph are given directions, such that the graph is acyclic. Once this has been done, it is possible to generate the corresponding schedule. The serialization function will assign operations to machines, and calculate the start time for each operation. The makespan can be determined after this process. It is important that the graph representation combined with the serialization function spans the entire set of feasible schedules. Furthermore, the serialization function should be deterministic in order to allow the local search to do its work.

To do this, we place the unscheduled operations in ordered queues for each machine type $T_k$. From these $t$ queues, we take the first operation that has no unscheduled job predecessor, and schedule it greedily into the first available machine, taking care to schedule it at the same start time or after that of the previously scheduled operation for that type. This ensures the ordering on the operations is preserved in the schedule, and conforms to the requirements that were laid out earlier in the representation section. Scheduling operations before the starting time of a previously scheduled operation on the same machine type can lead to a schedule that is actually infeasible.

@@

We could choose to apply several tricks to generate a better schedule. Instead, we will keep the serialization as simple as possible, and allow the local search do it for us by modifying the order of the operations for each machine type.

@@

\[ T_1\colon O_1 \rightarrow O_8 \]
\[ T_2\colon O_2 \rightarrow O_6 \rightarrow O_4 \]
\[ T_3\colon O_7 \rightarrow O_5 \rightarrow O_3 \]

@@

\subsubsection{Waiting cycle}

The acyclic property of the directed graph is crucial to the serialization process. It determines whether a configuration results in a feasible schedule or not. Suppose, using the above example, the arcs were modified to produce the following different queues for $T_1$ and $T_3$:

\[ T_1\colon O_8 \rightarrow O_1 \]
\[ T_3\colon O_5 \rightarrow O_3 \rightarrow O_7 \]

In this case, a cycle exists from $O_1 \rightarrow O_3 \rightarrow O_7 \rightarrow O_8$, and back to $O_1$. When the scheduler tries to serialize this configuration, it schedules $O_5$ as normal. Then it has $O_8$ to schedule on $T_1$, and $O_3$ on $T_3$. However, $O_8$ cannot be scheduled until $O_7$ is scheduled, and $O_3$ cannot be scheduled until $O_1$ (and $O_2$) is sceduled. This will be referred to as a \emph{waiting cycle}.

@@

For the FJS, we can define the longest path in a similar manner. Whereas the longest path for the classical Job Shop problem was also chronologically sequential in terms of the generated schedule, this is no longer necessarily the case for the FJS. We must also consider operations on the same machine type, scheduled on different machines. For example, consider the operations $O_{1,1}$ and $O_{2,2}$ from different jobs that must be scheduled on machines of type $T_k$, connected by a directed edge from $O_{1,1}$ to $O_{2,2}$. Suppose $O_{1,1}$ has been scheduled at time $t_1$ on machine $M_{k,1}$ due to a precedence constraint. Further suppose that machine $M_{k,2}$ is available from some time $t_2 < t_1$. However, the serialization function is forced to schedule $O_{2,2}$ on $M_{k,2}$ at time $t_1$ anyway. In this scenario, we would like to reverse the order of the arc between $O_{1,1}$ and $O_{2,2}$, so that $O_{2,2}$ may be scheduled earlier and therefore potentially decrease the length of the longest path.

@@

Let's consider the ordered queues containing the set of operations on a machine type, as defined by the acyclic directed graph corresponding to a feasible schedule. In the classical Job Shop case, a critical block swap identifies two consecutively scheduled nodes on a machine. These two nodes are necessarily consecutive in the queue, as well. In the FJS problem, this is no longer necessarily the case, since operations in between may have been scheduled on other machines of the same type. If one of these other operations is of the same job as either of the selected nodes, simply reversing the directed edge between the two nodes will lead to one or more cycles.

@@

	\ganttgroup[inline=false]{$M_{1,1}$}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,12}{3}{4}
	\ganttbar[bar/.style={fill=blue}]{1,10}{6}{7} \\

	\ganttgroup[inline=false]{$M_{1,2}$}{0}{1}
	\ganttbar[bar/.style={fill=blue}]{1,9}{2}{5}
	\ganttbar[bar/.style={fill=yellow}]{2,13}{6}{8} \\

	\ganttgroup[inline=false]{$M_{2,1}$}{0}{1}
	\ganttbar[bar/.style={fill=blue}]{1,8}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,14}{9}{11} \\

	\ganttgroup[inline=false]{$M_{3,1}$}{0}{1}
	\ganttbar[bar/.style={fill=yellow}]{2,11}{0}{2}

@@

Consider the schedule shown in \textbf{Figure \ref{fig:sch-swap}}. The ordered queue for $T_1$ that generated this schedule follows:

@@

We wish to swap the order of operations $O_{1,9}$ and $O_{2,13}$, both on the critical path. Simply swapping their places leads to cycles, since the job operations will be out of order. We can resolve this situation by reversing a few more arcs, to achieve the following ordering.

@@

Suppose the method above was not able to swap two operations, $O_{i,a_1}$ and $O_{j,b_2}$ for $i \neq j$. Then there exist operations $O_{i,a_2}$ and $O_{j,b_1}$ with $a_1 < a_2$ and $b_1 < b_2$ such that the ordered queue is equivalent to the following:

@@

The critical path neighborhood for the classical Job Shop is guaranteed to consist of only feasible schedules (acyclic graphs), a very nice result from Van Laarhoven et al \cite{simulatedannealingjsp}. Unfortunately, this is not the case for the Flexible Job Shop. We have slightly modified the swap operation to ensure no waiting cycles within a machine type are created. However, this does not guarantee no waiting cycles in the entire graph are generated. Consider the schedule shown in \textbf{Figure \ref{fig:sch-cycle}}, where colors correspond to jobs.

@@

	\ganttgroup[inline=false]{$M_{1,1}$}{656}{657}
	\ganttbar[bar/.style={fill=blue}]{1225}{656}{664}
	\ganttbar[bar/.style={fill=green}]{1298}{665}{668} \\

	\ganttgroup[inline=false]{$M_{1,2}$}{656}{657}
	\ganttbar[bar/.style={fill=yellow}]{905}{657}{665} \\

	\ganttgroup[inline=false]{$M_{1,3}$}{656}{657}
	\ganttbar[bar/.style={fill=darkblue}]{1009}{657}{658}
	\ganttbar[bar/.style={fill=red}]{1055}{659}{661}
	\ganttbar[bar/.style={fill=orange}]{1249}{662}{668} \\

@@

	\ganttgroup[inline=false]{$M_{4,1}$}{656}{657}
	\ganttbar[bar/.style={fill=darkblue}]{1010}{660}{663} \\

	\ganttgroup[inline=false]{$M_{4,4}$}{656}{657}
	\ganttbar[bar/.style={fill=green}]{1297}{660}{660}

@@

The acyclic directed graph that generated this schedule could have the following partial ordered queue for machine type $T_1$:

@@

Unfortunately, given the more restrictive formulation, there are no benchmark instances or results from the literature to compare to. Classical Job Shop instances do fit in our formulation, so a few have been included just for reference. These instances, \emph{ft06}, \emph{ft10}, and \emph{ft20} were generated by Fisher and Thompson (available on the web \cite{jobshopinfo}).

The next set of instances come from a previous project on exact methods for the FJS \cite{cijvat}, therefore including the optimal values. These instances are labeled \emph{c1 -- c19}. It should be noted that in almost all of these instances, the optimal makespan is equal to the maximum job makespan. This makes it easy to prove optimality, and also much easier to find an optimum.

The rest of the instances were generated using a random instance generator built for this purpose. These are labeled \emph{v10 -- v43}. The optimal makespans for these instances is unknown. However, they do provide insight into the convergence behavior, especially when compared between instances.

The local search was implemented in C, using the GCC compiler with the -O3 flag. The times listed are for an Intel Core i3 M370 @ 2.40 Hz. These results can be examined in \textbf{Table \ref{table:fjs}}.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{\bfseries}rrrrrrrrrrrrr}
	\toprule
	\multicolumn{4}{c}{Instance} && \multicolumn{2}{c}{Properties} && \multicolumn{5}{c}{Results} \\
	\cmidrule(r){1-4} \cmidrule(lr){6-7} \cmidrule(lr){9-13}
	\normalfont{name} & $|\mathcal{J}|$ & $|\mathcal{T}|$ & $|\mathcal{M}|$ && $\max C_\text{job}$ &
	$C_\text{opt}$ && $\alpha$ & restarts & $\bar{C}$ & $C_\text{best}$ & $\bar{t}$ \\
	\midrule
	ft06 &  6 &  6 &  6 &&   47 &   55 &&   0.99 & 1 &   55 &   55 &    0 \\
	ft10 & 10 & 10 & 10 &&  655 &  930 &&    0.9 & 5 &  954 &  941 &    2 \\
	     &    &    &    &&      &      &&   0.99 & 5 &  943 &  938 &   18 \\
	     &    &    &    &&      &      &&  0.999 & 5 &  933 &  930 &  182 \\
	ft20 & 20 &  5 &  5 &&  387 & 1165 &&    0.9 & 5 & 1215 & 1186 &    1 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 1182 & 1178 &   13 \\
	     &    &    &    &&      &      &&  0.999 & 5 & 1170 & 1165 &  130 \\[2ex]
	c1   &  4 &  2 &  4 &&   34 &   34 &&    0.9 & 1 &   34 &   34 &    0 \\
	c2   &  4 &  2 &  4 &&   53 &   53 &&    0.9 & 1 &   53 &   53 &    0 \\
	c3   &  4 &  2 &  4 &&  133 &  143 &&    0.9 & 1 &  143 &  143 &    0 \\
	c4   &  6 &  2 &  4 &&   52 &   57 &&   0.99 & 5 &   57 &   57 &    2 \\
	c5   &  6 &  2 &  4 &&  105 &  113 &&   0.99 & 5 &  113 &  113 &    2 \\
	c6   &  6 &  2 &  4 &&  205 &  225 &&   0.99 & 5 &  225 &  225 &    2 \\
	c7   &  4 &  3 &  9 &&   39 &   39 &&    0.9 & 1 &   39 &   39 &    0 \\
	c8   &  4 &  3 &  9 &&   64 &   64 &&    0.9 & 1 &   64 &   64 &    0 \\
	c9   &  4 &  3 &  9 &&  138 &  138 &&    0.9 & 1 &  138 &  138 &    0 \\
	c10  &  6 &  3 &  9 &&   50 &   50 &&    0.9 & 1 &   50 &   50 &    0 \\
	c11  &  6 &  3 &  9 &&  103 &  103 &&    0.9 & 1 &  103 &  103 &    0 \\
	c12  &  6 &  3 &  9 &&  204 &  204 &&    0.9 & 1 &  204 &  204 &    0 \\
	c13  &  8 &  3 &  9 &&  337 &  337 &&    0.9 & 5 &  337 &  337 &    0 \\
	c14  &  8 &  3 &  9 &&  705 &  705 &&    0.9 & 5 &  705 &  705 &    0 \\
	c15  &  8 &  3 &  9 && 1458 & 1458 &&    0.9 & 5 & 1458 & 1458 &    0 \\
	c16  & 10 &  4 & 16 &&  405 &  405 &&    0.9 & 5 &  405 &  405 &    0 \\
	c17  & 10 &  4 & 16 &&  955 &  955 &&    0.9 & 5 &  955 &  955 &    0 \\
	c18  & 10 &  4 & 16 && 1621 & 1621 &&    0.9 & 5 & 1621 & 1621 &    0 \\
	c19  & 15 &  5 & 25 && 2631 & 2631 &&    0.9 & 5 & 2631 & 2631 &    0 \\[2ex]
	v20  &  8 &  2 &  4 &&      &      &&    0.9 & 5 &      &      &      \\
	v30  & 16 &  2 &  4 &&  666 &      &&    0.9 & 5 & 2382 & 2381 &   10 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 2381 & 2381 &  102 \\
	v31  & 16 &  2 &  8 &&  566 &      &&    0.9 & 5 &  963 &  961 &   12 \\
	     &    &    &    &&      &      &&   0.99 & 5 &  960 &  959 &  126 \\
	v32  & 16 &  4 &  8 &&  655 &      &&    0.9 & 5 & 1119 & 1116 &   11 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 1112 & 1111 &  116 \\
	\bottomrule
\end{tabular}
\caption{Computational results for the FJS}
\label{table:fjs}
\end{table}

Most of the parameters in the table are self-explanatory. $\alpha$ is the cooling parameter. It is responsible for lowering the tempterature of the simulated annealing process. A value closer to 1 means the temperature cools slower, and thus the heuristic searches longer. The restarts column indicates how many times the temperature was reset. This entire process was repeated five times for every instance, to provide an average best solution found ($\bar{C}$) and the average running time ($\bar{t}$).

\subsection{Analysis}

For the Job Shop instances (\emph{ft06}, \emph{ft10}, and \emph{ft20}), the code as well as previous results in literature \cite{simulatedannealingjsp}. This gives us confidence that the code is robust.

The series of instances from the previous project (\emph{c1 -- c19}) are largely uninteresting, due to there being too few jobs in relation to machines. Interestingly, the few instances whose optimal makespans were larger than the lower bound given by the job makespans gave the search more trouble, despite being much smaller than some of the other instances. In several cases, the search failed to find an optimal solution despite the restarts, being very slightly off. However, this was averaged out by the other results. It should be noted that there were several instances of the algorithm becoming stuck, with the only neighborhood solution available being infeasible. This is very rare, but it highlights the importance of random restarts, when using the critical path neighborhood for the FJS.

For the other instances, we see that the algorithm converges very quickly. More searching failed to find solutions significantly below the best solutions found almost immediately. This indicates the critical path neighborhood works very well for the Flexible Job Shop problem. As the solution improves, the size of the neighborhood becomes smaller and smaller. This leads to very fast converging behavior.

It is also interesting to note that the amount of machine types and total machines have little to do with the "difficulty" of the instance. The algorithm converged equally quickly for all machine type configurations, with the amount of jobs and operations held constant.

\subsection{Serialization}

Intuitively, we can use the exact same representation and critical path neighborhood for the FBJS problem as for the FJS. The serialization function will have to be modified slightly to account for the blocking constraints. This is relatively simple -- the greedy algorithm will simply mark a machine as blocked after scheduling an operation, and unmark it when its successor is scheduled.

As mentioned in the introduction, the blocking constraint introduces new types of cycles.

\subsubsection{Pure blocking cycle}

An example of a pure blocking cycle is one where operation $O_{j_1,a}$ has been scheduled on machine $M_{t_1,m}$ and $O_{j_2,b}$ has been scheduled on $M_{t_2,n}$. Both of these machines are now blocked. Suppose there are no other available machines for machine types $T_{t_1}$ and $T_{t_2}$, and the scheduler now encounters $O_{j_2,b+1}$ to be scheduled on type $T_{t_1}$ and $O_{j_1,a+1}$ for $T_{t_2}$. Since the only possible machines for these types are blocked, this can only be resolved if the two machines can swap jobs simultaneously, as described in the introduction. Since we allow for blocking with swaps, the scheduler must recognize this type of cycle and schedule the two (or more) operations simultaneously.

Unfortunately, this cycle does not show up in our disjunctive graph representation, so we cannot detect it using traditional graph algorithms. We can choose to extend our representation, and indeed this may be the most complete choice. However, the scheduler will encounter maximally 1 operation to be scheduled on each machine type at once. Thus these cycles are guaranteed to be small and quite simple. By running through the blocked operations in $O(n^3)$ time, the scheduler can reliably identify these cycles and resolve them.

\subsubsection{Variable cycle}

A variable cycle is a mix of waiting and blocking. For example, suppose operation $O_{j_1,a}$ has been scheduled on machine $M_{T_i, k}$, the only available machine of its type. It is now blocked, and cannot schedule the next operation, $O_{j_2,b}$. However, this means $T_j$ cannot schedule its next operation, $O_{j_2,b + 1}$. $O_{j_1,a + 1}$ is unscheduled and ordered after $O_{j_2,b + 1}$, creating a deadlock. One machine is blocked, while the other is waiting.

This is an infeasible solution and cannot be resolved without reversing some edges. In this paper, we simply reverse the operation leading to this configuration, and try something else.

\subsection{Naive shift neighborhood}

There is little literature on the FBJS, and as a result there are few benchmark instances and results. In any case, these would not be applicable here due to the less generalized formulation we have used. To better assess the critical path neighborhood, we introduce a naive shift neigbhorhood for comparison.

In this neighborhood, we take a random operation of a random machine type, and shift it a random number of (permissable) places left or right. It may also take all of operations of the job this operation belongs to, and shift them all an individually random number of places left. The blocking constraint means that the operations belonging to a job must be executed as close together as possible. In this sense, it makes sense to move these operations as a unit, rather than individually.

While this neighborhood somewhat takes the nature of blocking and jobs into account, it doesn't use any information concerning the critical path, waiting and/or blocking times, or other useful properties. It is therefore referred to as a naive neighborhood.

@@

\subsection{Analysis}

\section{Conclusions and further research}

\printbibliography[heading=bibintoc]
